# Diabetes_Prediction_Dataset

```{r}
#install.packages("dplyr")

library("dplyr")

library(ggplot2)
```

```{r}
diabetes_prediction_dataset_csv_1_
head(diabetes_prediction_dataset_csv_1_)

```

### Use filter()

##### \*Filter the observations of your dataset based on one or more variable using filter

```{r}
## Filter for ages less then 40
filter(diabetes_prediction_dataset_csv_1_, age < 40)

## Filter for no info smoking history
filter(diabetes_prediction_dataset_csv_1_, is.na(smoking_history))

## Filter BMI
filter(diabetes_prediction_dataset_csv_1_, between(bmi, 50, 70))


```

### Use select()

#### \*Create a smaller dataset with a subset of variables using select()

```{r}

select(diabetes_prediction_dataset_csv_1_, age, smoking_history, bmi)

# select columns from age to bmi
select(diabetes_prediction_dataset_csv_1_, age:bmi)

select(diabetes_prediction_dataset_csv_1_, -c(gender, diabetes,HbA1c_level ))

```

### Use mutate()

#### \*Add two new columns to the data set

```{r}
#smaller data set 
diabetes_prediction_dataset_csv_1__small <- select
(diabetes_prediction_dataset_csv_1_, age, bmi, blood_glucose_level,HbA1c_level, diabetes)

diabetes_prediction_dataset_csv_1__small

# first mutate
mutate(diabetes_prediction_dataset_csv_1__small, Average_blood_glucose = (HbA1c_level * 35.6)-77.3)

# second mutate 
mutate(diabetes_prediction_dataset_csv_1__small, diabetes_age = mean(age, na.rm = T), .by = diabetes)


```

```         
```

### Use summarize()

##### \*create a data table of grouped summaries on at least one numeric variable by at least one categorical variable using summarize()

```{r}
summarize(diabetes_prediction_dataset_csv_1_,
          mean_age = mean(age, na.rm = T))

summarize(diabetes_prediction_dataset_csv_1_,
          mean_age = mean(age, na.rm = T),
          .by = gender)

summarize(diabetes_prediction_dataset_csv_1_,
          mean_bmi = mean(bmi, na.rm = T),
          count = n(),
          .by = c(diabetes))
```

#### Use arrange()

##### \*reorder a data table of your choice by one or more variables using arrange()

```{r}
arrange(diabetes_prediction_dataset_csv_1_, desc(bmi))

arrange(diabetes_prediction_dataset_csv_1_, bmi, diabetes)
```
